Online Variable Elimination for Credal Polytrees

### 2017-03-02 paper.tex
better abstract

We propose a variable-elimination algorithm for exact inference in singly connected credal networks with strong independence. Compared to earlier works in this direction, the necessary extensive specification of the local models and the inner points removal are performed in an online fashion. This might produce considerable savings in memory and, for the inner points removal only, time. Moreover, we show that, with singly connected networks, the inner points removal from the intermediate results of the inference can be achieved before the elimination step in a space with the same dimensionality of the variable to eliminate. This prevents issues related to convex hulling in high dimensional spaces. Despite the NP-hardness of the inference, even in the case of singly connected topologies, the experiments on a benchmark of randomly generated credal polytrees shows that exact inference can be achieved in models up to xxx variables if the number xxx. A further speed-up, at the price of an approximation, is achieved by rounding the values of the extreme points before the convexification, while an extension to general topologies can be achieved by loopset conditioning.

>>>>

We consier a variable-elimination procedure for exact inference in singly connected credal networks with strong independence. Compared to earlier works in this direction, the necessary extensive specification of the local models and the inner points removal are performed in an online fashion. This might produce considerable savings in memory and, for the inner points removal only, time. Moreover, we show that, with singly connected networks, the inner points removal from the intermediate results of the inference can be achieved before any elimination in a space with the same dimensionality of the variable to eliminate. This prevents computational issues related to convex hulling in high dimensional spaces. Despite the NP-hardness of the task, which concerns even singly connected topologies, the experiments on a benchmark of randomly generated credal polytrees shows that exact inference can be achieved in separately specified models up to xxx variables, xxx states, and indegree xxx. Faster approximate results can be achieved by simple rounding the values of the extreme points. Finally, multiply connected models can be also processed, with an outer approximation, with loopset conditioning.
